---
title: "BT4BR Assignment 7"
author: 
  name: Bartosz Kądziołka
  date: 05.12.2025
  tidy: false
format: 
  html:
      embed-resources: true
      theme: cerulean
editor: 
  markdown: 
    wrap: 72
---

## Assignment 7:

Use **tidyr** and **dplyr** libraries whenever possible, to do the
following:

**Task 1:** Load the "iris" dataset and transform it to longer.

-   tip: you should get 3 columns Species \| Trait \| Size

**Task 2:** Answer the following questions.

-   tip: you can answer each question "piping" only 2-3 commands

**Task 2.1:** What is the **mean** **length** and **width** of the
**sepal** and **petal** of each species?

**Task 2.2:** What is the **largest** and **widest**,  **petal** and
**sepal,**  of **virginica** and **versicolor** species?

**Task 2.3:** For each species, for sepal and petal width and length,
how many observations are above the mean?\*

-   tip: adding the means as column might help

**\***Might seem difficult (it is not that difficult if you know the
tricks ;) ). If you don't manage to make it work, submit your best
attempts and we will discuss it in class.

**Task 3:** Deliver a link to GitHub containing the a **Quarto** file
and the compiled **HTML**. It should contain explanations of what each
chunk does. If you have private repository, remember to send me
invitation so I can access it.

# **Task 1**

```{r}
library('tidyverse')
library('knitr')
data("iris")
force(iris)
long_iris <- pivot_longer(iris, cols = 1:4, names_to = "Trait", values_to = "Size")
```

Here I load the *tidyverse* library, then, I load the *iris* data set
and covert it to the long format with the `pivot_longer()` function. I
also load the knitr library for formatting my results later.

# Task 2

## Task 2.1

```{r}
#| message: false
mean_iris <- long_iris %>% 
group_by(Species, Trait) %>% 
summarise(MeanValue = mean(Size))
```

Here, I first group my values by *Species* and *Trait* before
calculating the mean lengths and widths and putting them into a new
table **mean_iris**.

## Task 2.2

```{r}
#| message: false
Max_Min_iris <- long_iris %>% 
filter(Species %in% c('virginica', 'versicolor')) %>% 
separate(col = Trait, into = c('Part', 'Measurement')) %>% 
group_by(Species, Part, Measurement) %>% 
summarise(Max = max(Size), Min = min(Size))
```

Now, I select only the **Virginica** and **Versicolor** species from the
*long_iris* table using the `filter()` function. Afterward, I separate
the *Trait* column and group the data by *Species*, *Part* and
*Measurement* before finding the max and min values inside the
`summarise()` function.

## Task 2.3

```{r}
#| message: false
counted_iris <- long_iris %>% 
full_join(mean_iris) %>% 
filter(Size > MeanValue) %>% 
group_by(Species, Trait) %>% 
count(name = 'Amount of reads above mean')
```

In this code chunk I start by joining the *long_iris* table, to the
*mean_iris* table with the standard condition which is
`join_by(Species, Trait)`. Then, I filter out the reads that are below
the mean value, leaving me with only the ones that are wider and longer
than the mean values. I then group the reads by *Species* and *Trait*,
finishing by counting the reads with the `count()` function.

# Results

```{r}
kable(head(long_iris))
kable(head(mean_iris,12))
kable(head(Max_Min_iris,8))
kable(head(counted_iris,12))
```
